Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2024-04-20 revision-1/lang.tex
checkpoint

  $$
  \begin{array}{c@{\hspace{5mm}}c}
  \begin{array}{rcl}
    \lcod{\store, v}{\cid} &=& v\\
    \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \fminus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fminus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}
  \end{array} & 
  \begin{array}{rcl}
    \lcod{\store, \flip{w}}{\cid} &=& \store(\elab{\flip{w}}{\cid})\\
    \lcod{\store, \secret{w}}{\cid} &=& \store(\elab{\secret{w}}{\cid})\\
    \lcod{\store, \mesg{w}}{\cid} &=& \store(\elab{\mesg{w}}{\cid})\\
    \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
    \lcod{\store, f\,e_1\,\cdots\, e_n}{\cid} &=& \delta(f,\lcod{\store, e_1}{\cid},\ldots,\lcod{\store,e_n}{\cid})
  \end{array}
  \end{array}
  $$

@@

\begin{fpfig}[t]{Top-to-bottom: Extended $\minifed$ syntax with BDOZ-style shares,
    interpretation of operations, and pre-processing properties of memories.}{fig-bdoz}
  {
$$
\begin{array}{rcl@{\hspace{8mm}}r}
  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,[m_1,\ldots,m_n])} &\\[2mm]
  %\textit{MACed\ values}\\[2mm]
  \be &::=& \cdots \mid \macv \mid \be \macplus \be \mid \be \mactimes \be \mid \be \macminus \be\\[2mm]
  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{\cid}
\end{array}
$$

@@

$$
\begin{array}{rcl}
  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \macplus (v^2, [m_1^2,\ldots,m^2_n])} &=&
  (v^1 \fplus v^2, [m_1^1 \fplus m_1^2 ,\ldots,m^1_n \fplus m_n^2])\\[2mm]
  \fcod{(v^1, [m_1^1,\ldots,m^1_n]) \mactimes v} &=&
   (v^1 \ftimes v, [m_1^1 \ftimes v ,\ldots,m^1_n \ftimes v])
\end{array}
$$
$$
\begin{array}{rcl}
  \lcod{\store, \be_1 \macplus \be_2}{\cid} &=&
  \fcod{\lcod{\store,\be_1}{\cid} \macplus \lcod{\store,\be_2}{\cid}}\\[2mm]  
  \lcod{\store, \be_1 \mactimes \be_2}{\cid} &=&
  \fcod{\lcod{\store,\be_1}{\cid} \mactimes \lcod{\store,\be_2}{\cid}}
\end{array}
$$

\begin{mathpar}
  \inferrule
      {\store(x) = v \\ \store(\macx{x}{\cid_1}) = (v_1,\ldots)\  \cdots\ 
        \store(\macx{x}{\cid_n}) = (v_n,\ldots)}
      {v = \fcod{v_1 \fplus \cdots \fplus v_n}}

  \inferrule
      {\store(\macx{x}{\cid'}) = (v,[m_1,\ldots,m_n]) \\  \cid \in [1..n]}
      {m_\cid = \fcod{\store(\mack{x}{\cid}) \fplus (\store(\elab{\flip{\ttt{delta}}}{\cid'}) \ftimes v)}}

@@

Eliminating cheating by adding information theoretically secure MAC
codes to shares is a fundamental approach accomplished by protocols
including BDOZ \cite{XXX} and SPDZ \cite{XXX}.  These protocols assume
pre-processing that distributes shares with MAC codes to all clients.
This integrates well with pre-processed Beaver Triples to implement
malicious secure and relatively efficient multiplication \cite{XXX}.
Here we show how to implement malicious secure two party multiplication
in the BDOZ style via some extensions of $\minifed$.

A field value $v$ is secret shared among $n$ clientz in BDOZ as follows.
Each client $\cid$ in a federation of size $n$ gets a value of the form
$(v_\cid,[m_1,\ldots,m_n])$, where $v_\cid$ is a secret
share of $v$, i.e., $v = \fcod{v_1 \fplus \cdots \fplus v_n}$. Further, 
values $[m_1,\ldots,m_n]$ are MACs of $v_\cid$, such that
$m_{\cid'} = k_{\cid'} + (k_\Delta * v_\cid)$ for each $\cid'\in [1..n]$,
where keys $k_{\cid'}$ and $k_\Delta$ are known only to $\cid$
and $k_\Delta$ is common to all MACs used by $\cid$. By
extending certain field operations to these forms, denoted $\macv$,
a form of semi-homorphic encryption is obtained that is adequate to compute
Beaver Triples. For example, if keys $k_1$ and $k_2$ authenticate
$\macv_1$ and $\macv_2$, then $k_1 \fplus k_2$ authenticates
$\macv_1 \macplus \macv_2$,  where $\macplus$ is addition
extended to MACed values. Relevant semantic rules for this, as
well as multiplication of MACed and field values, are given in
Figure \ref{fig-bdoz}- for more details the reader is referred to
\cite{XXX}.

We define variable forms $\macx{\secret{w}}{\cid}$ and
$\macx{\flip{w}}{\cid}$ range over local MACed shares of a secret and
random sample owned by client $\cid$, respectively. We let
$\mack{\macx{x}{\cid_1}}{\cid_2}$ denote the key that
authenticates $\cid_n$'s share of $\cid_1$'s value $x$.
We also assume a key $\flip{\ttt{delta}}$ on each client
$\cid$, which is that client's $k_\Delta$, 
In Figure \ref{fig-bdoz} we define the relevant properties
of pre-processed stores.

Similarly to field values, we can use MACed shares in
uniform independent distributions as one-time-pads for
hiding other values, using subtraction over MACed shares.
So we introduce a function $\macotp$ to do this
declaratively.
$$
\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
$$
Also, we can define a $\macauth$ function that authenticates
a client $\cid$'s share, and an $\macopen$ function
that extracts a field value from its shares.
\begin{mathpar}
\delta(\macauth, (v, [\ldots,m_\cid,\ldots]), k_\cid) \defeq
     (v, [\ldots,m_\cid,\ldots]) \text{\ if\ } m_\cid = \fcod{k_\cid \fplus (k_\Delta \ftimes v)}

\delta(\macopen, (v_1, [\ldots]), \ldots, (v_n, [\ldots]),) \defeq
\fcod{v_1 \fplus \cdots \fplus v_n}
\end{mathpar}
This authentication eliminates cheating and guarantees protocol
integrity. As we will show in Section \ref{section-types}, this
can be reflected in a security type for $\macauth$.

Putting these pieces together, we can implement 2-party multiplication with
Beaver Triples as follows. Recall that Beaver Triples are values
$(a,b,c)$ such that $c = \fcod{b \ftimes a}$- we implement them here
as a secret $\secret{c}$ and samples $\flip{a}$ and $\flip{b}$ owned
by an \emph{oracle} $\Oracle$ that we always assume is honest but is
otherwise not involved in the protocol. We define pre-processing
properties of Beaver Triples in Figure \ref{figure-bdoz}. First,
clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus \flip{a}$ and
$\secret{y} \fplus \flip{b}$ where
$\secret{x}$ and $\secret{y}$ are 1 and 2's secrets respectively.
Note that both perform a \emph{secure opening} by authenticating
each others' shares and record $\secret{x} \fplus \flip{a}$ and
$\secret{y} \fplus \flip{b}$ as $\mesg{d}$ and $\mesg{e}$ respectively.
$$
\begin{array}{lcl}
  \elab{\mesg{a}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{1}\\
  \elab{\mesg{a}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{2}\\
  \elab{\mesg{b}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{1}\\
  \elab{\mesg{b}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{2}\\
  \elab{\mesg{d}}{1} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
  \elab{\mesg{e}}{1}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
  \elab{\mesg{d}}{2} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
  \elab{\mesg{e}}{2}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
\end{array}
$$
Now, each client uses their share of $\secret{c}$ to generate their
share of the final output, which is again reported via a secure opening.
Homomorphic key calculations are elided for brevity.
$$
\begin{array}{lcl}
  \rvl{1} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{1}\\
  \rvl{2} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{2}\\
  \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \macauth(\rvl{2},\ldots))}{2} 
\end{array}
$$
Some information about input secrets is revealed by the product, but
secure opening ensures high integrity and hence robustness to
corruption, aka robust declassification.

>>>>

  $$
  \begin{array}{c@{\hspace{5mm}}c}
  \begin{array}{rcl}
    \lcod{\store, v}{\cid} &=& v\\
    \lcod{\store, \be_1 \fplus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fplus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \fminus \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \fminus \lcod{\store, \be_2}{\cid}}\\ 
    \lcod{\store, \be_1 \ftimes \be_2}{\cid} &=& \fcod{\lcod{\store, \be_1}{\cid} \ftimes \lcod{\store, \be_2}{\cid}}
  \end{array} & 
  \begin{array}{rcl}
    \lcod{\store, \flip{w}}{\cid} &=& \store(\elab{\flip{w}}{\cid})\\
    \lcod{\store, \secret{w}}{\cid} &=& \store(\elab{\secret{w}}{\cid})\\
    \lcod{\store, \mesg{w}}{\cid} &=& \store(\elab{\mesg{w}}{\cid})\\
    \lcod{\store, \rvl{w}}{\cid} &=& \store(\rvl{w})\\
    \lcod{\store, f\,\be_1\,\cdots\, \be_n}{\cid} &=& \delta(f,\lcod{\store, \be_1}{\cid},\ldots,\lcod{\store,\be_n}{\cid})
  \end{array}
  \end{array}
  $$

@@

\begin{fpfig}[t]{Top-to-bottom: Extended $\minifed$ syntax with BDOZ-style shares,
    interpretation of operations, and pre-processing properties of memories.}{fig-bdoz}
  {
$$
\begin{array}{rcl@{\hspace{8mm}}r}
  \multicolumn{3}{l}{m,k \in \mathbb{Z}_p \qquad \macv ::= (v,m)} &\\[2mm]
  %\textit{MACed\ values}\\[2mm]
  \be &::=& \cdots \mid \macv \mid \be \macplus \be \mid \be \mactimes \be \mid \be \macminus \be\\[2mm]
  x &::=& \cdots \mid \macx{\secret{w}}{\cid} \mid \macx{\flip{w}}{\cid} \mid \mack{x}{}
\end{array}
$$

\rule{130mm}{0.5pt}
\smallskip
$$
\begin{array}{rcl}
  \fcod{(v_1, m_1) \macplus (v_2, m_2)} &=& (v_1 \fplus v_2, m_1 \fplus m_2)\\[2mm]
  \fcod{(v, m) \mactimes v'} &=& (v \ftimes v', m \ftimes v')
\end{array}
$$
$$
\begin{array}{rcl}
  \lcod{\store, \be_1 \macplus \be_2}{\cid} &=&
  \fcod{\lcod{\store,\be_1}{\cid} \macplus \lcod{\store,\be_2}{\cid}}\\[2mm]  
  \lcod{\store, \be_1 \mactimes \be_2}{\cid} &=&
  \fcod{\lcod{\store,\be_1}{\cid} \mactimes \lcod{\store,\be_2}{\cid}}
\end{array}
$$

\medskip

@@

\begin{mathpar}
  \inferrule
      {\store(\macx{x}{\cid_1}) = (v_1,m_1) \\
        \store(\macx{x}{\cid_2}) = (v_2,m_2)}
      {\store(x) = \fcod{v_1 \fplus v_2}}

  \inferrule
      {\store(\elab{x}{\cid'}) = (v,m) \\ \cid \ne \cid'}
      {m = \lcod{\store, \mack{x}{} \fplus \flip{\ttt{delta}} \ftimes v}{\cid}}

@@

Eliminating cheating by adding information theoretic secure MAC codes
to shares is a fundamental approach realized by protocols such as BDOZ
\cite{XXX} and SPDZ \cite{XXX}.  These protocols assume a
pre-processing phase that distributes shares with MAC codes to
clients.  This integrates well with pre-processed Beaver Triples to
implement malicious secure, and relatively efficient, multiplication
\cite{XXX}.  Here we show how to implement malicious secure two party
multiplication in the BDOZ style via some extensions of $\minifed$.

A field value $v$ is secret shared among 2 clientz in BDOZ as follows.
Each client $\cid$ gets a pair of the form $(v_\cid,m_\cid)$, where
$v_\cid$ is a secret share of $v$, i.e., $v = \fcod{v_1 \fplus v_2}$.
Further, field values $m_\cid$ are MACs of $v_\cid$ that are authenticated
by the other client's key. In particular, $m_\cid = k + (k_\Delta *
v_\cid)$ where \emph{keys} $k$ and $k_\Delta$ are known only to $\cid'
\ne \cid$ and $k_\Delta$. Further, there is a unique \emph{local key}
$k$ for each shared value, while the \emph{global key} $k_\Delta$ is
common to all MACs authenticated by $\cid'$. By extending certain
field operations to these MACed values, a semi-homorphic encryption
scheme is obtained that is adequate to compute Beaver Triples. For
example, if local keys $k_1$ and $k_2$ authenticate $\macv_1$ and
$\macv_2$, then $k_1 \fplus k_2$ authenticates $\macv_1 \macplus
\macv_2$, where $\macplus$ is addition extended to MACed
values. Semantics for summation of MACed shares, and multipication of
shares and plain field values, are given in Figure \ref{fig-bdoz}.
For more details the reader is referred to \cite{XXX}. We note
that while we restrict values $v$, $m$, and $k$ to the same
field in this presentation for simplicity, in general $m$ and
$k$ can be in extensions of $\mathbb{Z}_p$.

We define variable forms $\macx{\secret{w}}{\cid}$ and
$\macx{\flip{w}}{\cid}$ to range over local MACed shares of a secret
and random sample owned by client $\cid$, respectively. We let
$\mack{x}{}$ denote the local key for authenticating shares of $x$.
We also assume a global key $\flip{\ttt{delta}}$ on each client.
Pre-processing is modeled by assuming that shares are provided on
input random tapes (memories) with the required properties, as defined
on the bottom of Figure \ref{fig-bdoz}.

Similarly to field values, we can use MACed shares in
uniform independent distributions as one-time-pads for
hiding other values, using subtraction over MACed shares.
So we introduce a function $\macotp$ to do this
declaratively.
$$
\delta(\macotp,v_1,v_2) \defeq \fcod{v_1 \macminus v_2}
$$

Another key idea in BDOZ is \emph{secure opening}, where distributed
shares of a value $v$ are communicated and combined to ``open'' the
same $v$ on each client. This relies on authentication to detect
cheating in the malicious setting. To implement secure opening we
define a $\macauth$ function that authenticates shares, and a
$\macopen$ function that extracts a field value from its shares.
\begin{mathpar}
\delta(\macauth, (v, m), k) \defeq
     (v, m) \text{\ if\ } m = k \fplus (\flip{\ttt{delta}} \ftimes v)

\delta(\macopen, (v_1, m_1), (v_2,m_2)) \defeq
\fcod{v_1 \fplus v_2}
\end{mathpar}
This authentication eliminates cheating and guarantees protocol
integrity. As we will show in Section \ref{section-types}, this
can be reflected in a security type for $\macauth$.

Putting these pieces together, we can implement 2-party multiplication with
Beaver Triples as follows. Recall that Beaver Triples are values
$(a,b,c)$ such that $c = \fcod{b \ftimes a}$- we implement them here
as a secret $\secret{c}$ and samples $\flip{a}$ and $\flip{b}$ owned
by an \emph{oracle} $\Oracle$ that we always assume is honest but is
otherwise not involved in the protocol. We define pre-processing
properties of Beaver Triples in Figure \ref{fig-bdoz}. Only one
Beaver triple can be used per multiplication to ensure secure
openings, but our scheme can be extended with more Beaver triples
and reuse of them will be rule out by type linearity as we will
discuss in Section \ref{section-types}.

First, clients 1 and 2 compute (aka \emph{open}) $\secret{x} \fplus
\flip{a}$ and $\secret{y} \fplus \flip{b}$ where $\secret{x}$ and
$\secret{y}$ are 1 and 2's secrets respectively.  Then they both
perform a secure opening of $\secret{x} \fplus \flip{a}$ and
$\secret{y} \fplus \flip{b}$, by authenticating each others' shares,
and record them as $\mesg{d}$ and $\mesg{e}$ respectively. Note
that the secrets remain hidden since $\flip{a}$ and $\flip{b}$
are random values.
$$
\begin{array}{lcl}
  \elab{\mesg{a}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{1}\\
  \elab{\mesg{a}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{a}}{\Oracle}}\ \macgv{\elab{\secret{x}}{1}})}{2}\\
  \elab{\mesg{b}}{2} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{1}\\
  \elab{\mesg{b}}{1} &:=&
  \elab{(\macotp\ \macgv{\elab{\flip{b}}{\Oracle}}\ \macgv{\elab{\secret{y}}{2}})}{2}\\
  \elab{\mesg{d}}{1} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{2} \fminus \mack{\macx{\flip{a}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{1}\\
  \elab{\mesg{e}}{1}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{2} \fminus \mack{\macx{\flip{b}}{\Oracle}}{2}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{1}\\
  \elab{\mesg{d}}{2} &:=&
  \elab{(\macopen\ \macauth(\mesg{a}, \mack{\macx{\secret{x}}{1}}{1} \fminus \mack{\macx{\flip{a}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{x}}{1}}\macminus\macgv{\elab{\flip{a}}{\Oracle}}))}{2}\\
  \elab{\mesg{e}}{2}&:=&
  \elab{(\macopen\ \macauth(\mesg{b}, \mack{\macx{\secret{y}}{2}}{1} \fminus \mack{\macx{\flip{b}}{\Oracle}}{1}) \ (\macgv{\elab{\secret{y}}{2}}\macminus\macgv{\elab{\flip{b}}{\Oracle}}))}{2}
\end{array}
$$
Now, each client uses their share of $\secret{c}$ to generate their
share of the final output, which is again reported via a secure opening.
Homomorphic key calculations are elided for brevity.
$$
\begin{array}{lcl}
  \rvl{1} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{1}\\
  \rvl{2} &:=&
  \elab{( (\mesg{d} \ftimes \mesg{e}) \macplus
          (\mesg{d} \mactimes \macgv{\elab{\flip{b}}{\Oracle}}) \macplus
          (\mesg{e} \mactimes \macgv{\elab{\flip{a}}{\Oracle}}) \macplus \macgv{\elab{\secret{c}}{\Oracle}}
    )}{2}\\
  \out{1} &:=& \elab{(\macopen\ \rvl{1} \ \macauth(\rvl{2},\ldots))}{1} \\
  \out{2} &:=& \elab{(\macopen\ \macauth(\rvl{1},\ldots) \ \rvl{2})}{2} 
\end{array}
$$
Some information about input secrets is revealed by the product, but
secure opening ensures high integrity and hence robustness to
corruption, aka robust declassification.

### 2024-04-20 revision-1/model.tex
checkpoint

>>>>

\subsection{Probability Mass Functions and Program Distributions}
\label{section-pmf}

Given a program $\prog$ with $\iov(\prog) = S \cup V$ and $\store \in
\mems(\iov(\prog))$, we are fundamentally concerned with the joint
probability of the assignments in $\store_V$ conditioned by the
assignments $\store_S$ in any given run, since probabilistic
dependence between honest secrets and corrupt views is our metric of
information leakage.

\subsection{Basic Definitions}

We begin by defining joint probability mass functions in the standard
manner, though following \cite{barthe2019probabilistic} we use
memories to denote mappings of variables to values (i.e., outcomes),
so for example given a pmf $\pmf$ we will write $\pmf(\{ \sx{1}{a}
\mapsto 0, \vx{2}{b} \mapsto 1 \})$ to denote the probability that
$\sx{1}{a} = 0$ and $\vx{2}{b} = 1$ in the final memory of a
program containing these variables.
\begin{definition}
  A \emph{probability mass function} $\pmf$ is a function
  mapping memories in $\mems(X)$ for some $X$ to values in $[0,1]$, such that:
  $$
  \sum_{\store \in \mems(X)} \pmf(\store) \  = \ 1
  $$
\end{definition}
%To recover succinct and familiar notation, we may omit the domain of a
%distribution when it is clear from an application context-
%i.e., we allow the following sugaring:
%$$
%\pdf{}(\store) \defeq \pdf{\dom(\store)}(\store)
%$$
Now, we can define a notion of marginal and conditional
distributions as follows, which are standard for discrete
probability mass functions. 
\begin{definition}
  Given $\pmf$ the \emph{marginal distribution} of variables $X$
  in $\pmf$, denoted $\margd{\pmf}{X}$, is defined as follows:
  $$
  \forall \store \in \mems(X) \quad . \quad \margd{\pmf}{X}(\store) =
  \sum_{\store' \in \mems(X-\dom(\dom(\pmf)))} \pmf(\store \cap \store')
  $$
\end{definition}

\begin{definition}
  Given $\margd{\pmf}{X}$, its \emph{conditional distribution given
  $\stores$} with $\dom(\stores) = Y$ with $Y \subseteq X$, denoted
  $\condd{\pmf}{X}{\stores}$, is a distribution with domain $X$ where for all
  $\store \in \mems(X)$:
  $$
  \condd{\pmf}{X}{\stores}(\store) =
  (\sum_{\store' \in \stores} \margd{\pmf}{X}(\store \cap \store')) /
  (\sum_{\store' \in \stores} \margd{\pmf}{Y}(\store'))
  $$
  where $\margd{\pmf}{X}(\store \cap \store')) = 0$ if $\store \cap \store'$ is undefined.
\end{definition}
To recover familiar notation we allow the syntactic
sugarings $\condd{\pmf}{X}{\store}  \defeq \condd{\pmf}{X}{\{ \store\}}$, and
$\pmf(\store)  \defeq \margd{\pmf}{X}(\store)$ and $\pmf(\store|\stores) \defeq
\condd{\pmf}{X}{\stores}(\store)$ where $\dom(\store) = X$.
%\begin{eqnarray*}
%  \condd{\pmf}{X}{\store}  &\defeq& \condd{\pmf}{X}{\{ \store\}}\\
%  \pmf(\store)  &\defeq& \margd{\pmf}{X}(\store)  \qquad \dom(\store) = X\\
%  \pmf(\store|\stores)  &\defeq& \condd{\pmf}{X}{\stores}(\store) \qquad \dom(\store) = X
%\end{eqnarray*}

\subsection{Program Distributions}
Now we can define the probability distribution of a program $\prog$,
that we denote $\progd(\prog)$. Since $\fedcat$ is deterministic the
results (view assignments) of any run are determined by the input
secrets together with the random tape. And since we constrain programs
to not overwrite views, we are assured that \emph{final} memories
contain both a complete record of all initial secrets as well as views
resulting from communicated information.

We consider random tapes to be selected from a uniform distribution.
In our threat model, we will also assume that any possible initial
secrets are equally likely when formulating the program distribution.
In this setting, given a program $\prog$ with $\iov(\prog) = S \cup
V$ and $\flips(\prog) = F$ we will consider all $\store
\in \mems(S \cup V \cup F)$ such that
$
\config{\store_{S \cup F}}{\prog} \redxs \config{\store_}{\varnothing}
$
to be equally probable, establishing the basic distribution of the
program. From this, we can immediately derive the marginal distribution
of $S \cup V$ to reason about dependencies between secrets and views. 
\begin{definition}
  \label{def-progd}
  \label{definition-progd}
  Given program $\prog$ with $\iov(\prog) = (S,V)$ and $\flips(\prog) = F$, define $\stores$ as:
  $$
  \stores \defeq \{ \store \mid (\dom(\store) = S \cup V \cup F) \wedge (\store_{S \cup F},\prog) \redxs (\store,\varnothing) \}
  $$
  We note immediately that $|\stores| = 2^{S \cup F}$ given the (deterministic) definition of $\redx$. 
  Define also $\pmf$ as the program's \emph{basic distribution} such that for all
  $\store \in \mems(S \cup V \cup F)$:
  $$
  \pmf(\store) =  1 / |\stores| \ \text{if}\ \store \in \stores, \text{otherwise}\ 0
  %\begin{cases}
  %  1 / |\stores| & \text{if}\ \store \in \stores\\
  %  0 & \text{otherwise}
  %\end{cases}
  $$
  Then the \emph{program distribution of $\prog$}, denoted $\progd(\prog)$, is the
  marginal distribution of $S \cup V$ in $\prog$'s basic distribution, i.e., 
  $
  \progd(\prog) \defeq  \margd{\pmf}{S\cup V}
  $.
  In some cases, we will also be concerned with the (joint)
  probabilities of expression interpretation given a preceding program
  execution, and we write $\progd(\prog, \be)$ to denote the program
  distribution $\progd(\prog;\itv := \be)$ where $\itv$ is a
  special variable that is never used in programs.
\end{definition}

\begin{definition}
  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{X}{Y}$ iff for all
    $\store \in \mems(X \cup Y)$ we have
  $\margd{\pmf}{X \cup Y}(\store) =
  \pmf(\store_X) * \pmf(\store_Y)$
\end{definition}
